---
title: "Bayes factor design analysis for daytime mEDI"
author: "Taisuke Eto"
date: "2026-07-30"
output: html_document
---

This script evaluates the planned sample size for the within-participant association between daytime melanopic EDI dose and melatonin suppression.
It simulates data, fits the Full and Null models, calculates Bayes factors, checks model diagnostics, and assesses random-slope recovery.

Daytime melanopic EDI dose is decomposed into:

- `edi_within`: the session-specific deviation from each participant's mean
- `edi_between`: the participant's mean exposure relative to the sample mean

The two fitted models retain the same between-participant fixed effect and random-effects structure and differ only in the population-level within-participant effect:

```text
Full:
y ~ edi_within + edi_between + (1 + edi_within | id)

Null:
y ~ edi_between + (1 + edi_within | id)
```

# 1. Load inputs and prepare the exposure predictor

The simulation uses parameter combinations created by `informed_data_simulation_daytimeEDI_sep.Rmd` and observed daytime mEDI profiles to preserve a realistic exposure distribution.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare-inputs}
library(future)
library(furrr)
library(brms)
library(bridgesampling)
library(MASS)
library(tidyverse)
library(progressr)
library(posterior)

# Prevent additional threads from being created inside each worker.
Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1"
)

# Reuse compiled Stan models when possible.
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = TRUE))

parallelly::availableCores()

informed_parameters <- read.csv("informed_parameters.csv") |>
  dplyr::select(
    intercept_mean,
    within_slope_mean,
    between_slope_mean
  ) |>
  distinct()

# Parallelisation remains at the informed-parameter-row level.
# Fewer workers are used automatically when the CSV contains fewer rows.

workers <- min(
  16,
  nrow(informed_parameters),
  max(1, parallel::detectCores(logical = TRUE) - 1)
)

plan(multisession, workers = workers)


# Load the standardised log-transformed daytime mEDI dose data
medi_dose_data <- read.csv("medi_data_log_transformed.csv") |>
  filter(
    is.finite(log_mean_dose)
  )

# Store each observed participant's exposure profile
medi_values_by_participant <- medi_dose_data |>
  group_by(Id) |>
  summarise(
    values = list(log_mean_dose),
    .groups = "drop"
  ) |>
  filter(lengths(values) > 0) |>
  pull(values)

# Sample complete participant profiles and then perform
# the within/between decomposition
sample_medi_profile <- function(n_participants, n_sessions) {

  donor_index <- sample.int(
    length(medi_values_by_participant),
    size = n_participants,
    replace = TRUE
  )

  sampled_values <- purrr::map(
    donor_index,
    function(index) {

      donor_values <- medi_values_by_participant[[index]]

      donor_values[
        sample.int(
          length(donor_values),
          size = n_sessions,
          replace = length(donor_values) < n_sessions
        )
      ]
    }
  )

  sampled_data <- tibble(
    id = factor(
      rep(seq_len(n_participants), each = n_sessions)
    ),
    session = rep(
      seq_len(n_sessions),
      times = n_participants
    ),
    edi_total = unlist(
      sampled_values,
      use.names = FALSE
    )
  ) |>
    group_by(id) |>
    mutate(
      # Participant mean based on the three simulated sessions
      edi_between_raw = mean(edi_total),

      # Session-specific deviation from the participant mean
      edi_within_raw = edi_total - edi_between_raw
    ) |>
    ungroup()

  # Use one between-person value per participant when calculating its SD
  between_values <- sampled_data |>
    distinct(id, edi_between_raw) |>
    pull(edi_between_raw)

  between_mean <- mean(between_values)
  between_sd <- sd(between_values)
  within_sd <- sd(sampled_data$edi_within_raw)

  if (
    !is.finite(between_sd) ||
    !is.finite(within_sd) ||
    between_sd == 0 ||
    within_sd == 0
  ) {
    stop(
      "Unable to standardise the sampled within/between mEDI components."
    )
  }

  sampled_data |>
    mutate(
      edi_between =
        (edi_between_raw - between_mean) / between_sd,

      edi_within =
        edi_within_raw / within_sd,

      id_numeric = as.integer(id)
    )
}



test_predictor_data <- sample_medi_profile(
  n_participants = 50,
  n_sessions = 3
)

# Within-person means must be approximately zero
test_predictor_data |>
  group_by(id) |>
  summarise(
    mean_within = mean(edi_within),
    .groups = "drop"
  ) |>
  summarise(
    maximum_absolute_mean =
      max(abs(mean_within))
  )

# Overall within component
test_predictor_data |>
  summarise(
    within_mean = mean(edi_within),
    within_sd = sd(edi_within)
  )

# Between component: one value per participant
test_predictor_data |>
  distinct(id, edi_between) |>
  summarise(
    between_mean = mean(edi_between),
    between_sd = sd(edi_between)
  )

```

# 2. Define simulation settings and priors

The main settings include the sample size, number of sessions, generating standard deviations, Bayes-factor threshold, MCMC settings, and diagnostic criteria.

```{r simulation-settings}
# Fixed participant number, dictated by resource limitations
n_ids <- 50

# Number of observed melatonin suppression outcomes per participant
n_sessions <- 3

# Standard deviations used to generate the data
intercept_sd <- 0.2
slope_sd <- 0.2
residual_sd <- 0.6

# Correlation between participant-specific intercepts and slopes
intercept_slope_cor <- 0

# Number of the simulation
num_simulations <- 100

# Define threshold for Bayes Factor
bf_threshold <- 10

# brms sampling settings
chains <- 4
iter <- 4000
warmup <- 1000

# Refit a failed model once with a longer MCMC run.
retry_failed_fits <- TRUE
retry_iter <- 8000
retry_warmup <- 2000

# Model-diagnostic thresholds
rhat_limit <- 1.01
ess_limit <- 100 * chains
bfmi_limit <- 0.30

# Criteria for random-slope recovery with three observations per participant
recovery_min_valid_rate <- 0.95
recovery_max_abs_relative_bias <- 0.20
recovery_min_coverage <- 0.90
recovery_max_coverage <- 1.00
recovery_max_collapse_rate <- 0.05
recovery_max_median_interval_width_ratio <- 2.00

# Treat an estimate below 25% of the true value as collapsed toward zero.
slope_sd_collapse_threshold <- 0.25 * slope_sd

# Proper priors are required for bridge sampling.
full_priors <- c(
  prior(
    normal(0, 1),
    class = "Intercept"
  ),
  prior(
    student_t(1, 0, sqrt(2) / 4),
    class = "b",
    coef = "edi_within"
  ),
  prior(
    student_t(1, 0, sqrt(2) / 4),
    class = "b",
    coef = "edi_between"
  ),
  prior(
    student_t(3, 0, 0.5),
    class = "sd"
  ),
  prior(
    lkj(2),
    class = "cor"
  ),
  prior(
    student_t(3, 0, 0.5),
    class = "sigma"
  )
)

null_priors <- c(
  prior(
    normal(0, 1),
    class = "Intercept"
  ),
  prior(
    student_t(1, 0, sqrt(2) / 4),
    class = "b",
    coef = "edi_between"
  ),
  prior(
    student_t(3, 0, 0.5),
    class = "sd"
  ),
  prior(
    lkj(2),
    class = "cor"
  ),
  prior(
    student_t(3, 0, 0.5),
    class = "sigma"
  )
)

# Random-intercept and random-slope covariance matrix
random_effect_covariance <- matrix(
  c(
    intercept_sd^2,
    intercept_slope_cor * intercept_sd * slope_sd,
    intercept_slope_cor * intercept_sd * slope_sd,
    slope_sd^2
  ),
  nrow = 2
)

```

# 3. Define the model-diagnostic function

This function checks HMC diagnostics, R-hat, effective sample size, and BFMI. 
It also extracts the random-slope standard deviation for the recovery assessment.

```{r diagnostic-function}

diagnose_brms_fit <- function(
    fit,
    chains,
    max_treedepth,
    true_slope_sd,
    slope_sd_collapse_threshold,
    rhat_limit = 1.01,
    ess_limit = 100 * chains,
    bfmi_limit = 0.30
) {

  # HMC diagnostics
  sampler_parameters <- rstan::get_sampler_params(
    fit$fit,
    inc_warmup = FALSE
  )

  n_divergent <- sum(
    vapply(
      sampler_parameters,
      function(x) sum(x[, "divergent__"]),
      numeric(1)
    )
  )

  n_max_treedepth <- sum(
    vapply(
      sampler_parameters,
      function(x) sum(x[, "treedepth__"] >= max_treedepth),
      numeric(1)
    )
  )

  bfmi_values <- rstan::get_bfmi(fit$fit)
  min_bfmi <- min(bfmi_values)

  # R-hat and ESS
  posterior_draws <- posterior::as_draws_df(fit)

  parameter_names <- posterior::variables(posterior_draws)

  # Exclude Stan sampler variables such as lp__, energy__, etc.
  model_parameter_names <- parameter_names[
    grepl(
      "^(b_|sd_|cor_|sigma$)",
      parameter_names
    )
  ]

  diagnostic_summary <- posterior::summarise_draws(
    posterior::subset_draws(
      posterior_draws,
      variable = model_parameter_names
    ),
    "rhat",
    "ess_bulk",
    "ess_tail"
  )

  all_diagnostics_finite <-
    all(is.finite(diagnostic_summary$rhat)) &&
    all(is.finite(diagnostic_summary$ess_bulk)) &&
    all(is.finite(diagnostic_summary$ess_tail))

  max_rhat <- max(
    diagnostic_summary$rhat,
    na.rm = TRUE
  )

  min_bulk_ess <- min(
    diagnostic_summary$ess_bulk,
    na.rm = TRUE
  )

  min_tail_ess <- min(
    diagnostic_summary$ess_tail,
    na.rm = TRUE
  )

  # Random-slope SD posterior
  slope_sd_draws <-
    posterior_draws[["sd_id__edi_within"]]

  estimated_slope_sd <- mean(slope_sd_draws)

  median_estimated_slope_sd <- median(slope_sd_draws)

  slope_sd_lower <- unname(
    quantile(slope_sd_draws, 0.025)
  )

  slope_sd_upper <- unname(
    quantile(slope_sd_draws, 0.975)
  )

  slope_sd_coverage <-
    slope_sd_lower <= true_slope_sd &&
    slope_sd_upper >= true_slope_sd

  slope_sd_interval_width <-
    slope_sd_upper - slope_sd_lower

  slope_sd_error <-
    estimated_slope_sd - true_slope_sd

  slope_sd_collapsed <-
    estimated_slope_sd < slope_sd_collapse_threshold

  # Strict diagnostic pass
  divergence_pass <-
    n_divergent == 0
  
  rhat_pass <-
    all_diagnostics_finite &&
    max_rhat <= rhat_limit
  
  bulk_ess_pass <-
    all_diagnostics_finite &&
    min_bulk_ess >= ess_limit
  
  tail_ess_pass <-
    all_diagnostics_finite &&
    min_tail_ess >= ess_limit
  
  bfmi_pass <-
    is.finite(min_bfmi) &&
    min_bfmi >= bfmi_limit
  
  treedepth_pass <-
    n_max_treedepth == 0
  
  diagnostic_pass <-
    divergence_pass &&
    rhat_pass &&
    bulk_ess_pass &&
    tail_ess_pass &&
    bfmi_pass &&
    treedepth_pass

  tibble::tibble(
    n_divergent = n_divergent,
    n_max_treedepth = n_max_treedepth,
    max_rhat = max_rhat,
    min_bulk_ess = min_bulk_ess,
    min_tail_ess = min_tail_ess,
    min_bfmi = min_bfmi,
    estimated_slope_sd = estimated_slope_sd,
    median_estimated_slope_sd = median_estimated_slope_sd,
    slope_sd_lower = slope_sd_lower,
    slope_sd_upper = slope_sd_upper,
    slope_sd_coverage = slope_sd_coverage,
    slope_sd_interval_width = slope_sd_interval_width,
    slope_sd_error = slope_sd_error,
    slope_sd_squared_error = slope_sd_error^2,
    slope_sd_collapsed = slope_sd_collapsed,
    diagnostic_pass = diagnostic_pass,
    divergence_pass = divergence_pass,
    rhat_pass = rhat_pass,
    bulk_ess_pass = bulk_ess_pass,
    tail_ess_pass = tail_ess_pass,
    bfmi_pass = bfmi_pass,
    treedepth_pass = treedepth_pass
  )
}
```

# 4. Run the BFDA simulations

For every simulated dataset, the full and null models are fitted with `brms`.
Their marginal likelihoods are estimated by bridge sampling. 
The resulting Bayes factor is \(BF_{10}\), in favour of the full model over the null model.

```{r run-bfda}
start_time <- Sys.time()

handlers("progress")

with_progress({

  p <- progressor(steps = nrow(informed_parameters) * num_simulations)

  bfda_simulated_data_parallel <- future_map_dfr(
    1:nrow(informed_parameters),
    function(row) {

      bf_list <- rep(NA_real_, num_simulations)
      log_bf_list <- rep(NA_real_, num_simulations)
      beta_posterior_mean_list <- rep(NA_real_, num_simulations)
      
      full_diagnostic_list <- vector(
        "list",
        num_simulations
      )
      
      null_diagnostic_list <- vector(
        "list",
        num_simulations
      )
      
      bridge_logbf_difference_list <-
        rep(NA_real_, num_simulations)
      
      bridge_decision_stable_list <-
        rep(NA, num_simulations)

      full_retry_list <- rep(FALSE, num_simulations)
      null_retry_list <- rep(FALSE, num_simulations)
      full_initial_diagnostic_pass_list <- rep(NA, num_simulations)
      null_initial_diagnostic_pass_list <- rep(NA, num_simulations)
      

      # Each worker compiles the Full and Null models only on its first simulation. Later simulations reuse these compiled models.
      fit_full <- NULL
      fit_null <- NULL

      for (simulation in seq_len(num_simulations)) {

        # Participant-specific random intercepts and slopes
        random_effects <- MASS::mvrnorm(
          n = n_ids,
          mu = c(0, 0),
          Sigma = random_effect_covariance
        )

        random_intercept <- random_effects[, 1]
        random_slope <- random_effects[, 2]

        # Population-level data-generating parameters
        intercept_mean <- 
          informed_parameters$intercept_mean[row]
        
        within_slope_mean <-
          informed_parameters$within_slope_mean[row]
        
        between_slope_mean <-
          informed_parameters$between_slope_mean[row]
        
        # Sample participant exposure profiles and create within- and between-participant predictors
        sim_data <- sample_medi_profile(
          n_participants = n_ids,
          n_sessions = n_sessions
        )
        
        # Generate the standardised melatonin suppression outcome
        sim_data$y <-
          intercept_mean +
          random_intercept[sim_data$id_numeric] +
          between_slope_mean * sim_data$edi_between +
          (
            within_slope_mean +
              random_slope[sim_data$id_numeric]
          ) * sim_data$edi_within +
          rnorm(
            nrow(sim_data),
            mean = 0,
            sd = residual_sd
          )

        if (anyNA(sim_data)) {
          stop(
            sprintf(
              "NA found in sim_data at row %d, simulation %d",
              row,
              simulation
            )
          )
        }

        if (is.null(fit_full)) {
          fit_full <- brm(
            y ~ edi_within + edi_between + 
              (1 + edi_within | id),
            data = sim_data,
            family = gaussian(),
            prior = full_priors,
            backend = "rstan",
            chains = chains,
            cores = 1,
            iter = iter,
            warmup = warmup,
            seed = 20260405 + row * 100000 + simulation,
            save_pars = save_pars(all = TRUE),
            control = list(
              adapt_delta = 0.95,
              max_treedepth = 12
            ),
            silent = 2,
            refresh = 0
          )
        } else {
          fit_full <- update(
            fit_full,
            newdata = sim_data,
            recompile = FALSE,
            chains = chains,
            cores = 1,
            iter = iter,
            warmup = warmup,
            seed = 20260405 + row * 100000 + simulation,
            control = list(
              adapt_delta = 0.95,
              max_treedepth = 12
            ),
            silent = 2,
            refresh = 0
          )
        }

        if (is.null(fit_null)) {
          fit_null <- brm(
            y ~ edi_between + 
              (1 + edi_within | id),
            data = sim_data,
            family = gaussian(),
            prior = null_priors,
            backend = "rstan",
            chains = chains,
            cores = 1,
            iter = iter,
            warmup = warmup,
            seed = 30260405 + row * 100000 + simulation,
            save_pars = save_pars(all = TRUE),
            control = list(
              adapt_delta = 0.95,
              max_treedepth = 12
            ),
            silent = 2,
            refresh = 0
          )
        } else {
          fit_null <- update(
            fit_null,
            newdata = sim_data,
            recompile = FALSE,
            chains = chains,
            cores = 1,
            iter = iter,
            warmup = warmup,
            seed = 30260405 + row * 100000 + simulation,
            control = list(
              adapt_delta = 0.95,
              max_treedepth = 12
            ),
            silent = 2,
            refresh = 0
          )
        }

        # Diagnose both fitted models
        full_diagnostics <- diagnose_brms_fit(
          fit = fit_full,
          chains = chains,
          max_treedepth = 12,
          true_slope_sd = slope_sd,
          slope_sd_collapse_threshold = slope_sd_collapse_threshold,
          rhat_limit = rhat_limit,
          ess_limit = ess_limit,
          bfmi_limit = bfmi_limit
        )
        
        null_diagnostics <- diagnose_brms_fit(
          fit = fit_null,
          chains = chains,
          max_treedepth = 12,
          true_slope_sd = slope_sd,
          slope_sd_collapse_threshold = slope_sd_collapse_threshold,
          rhat_limit = rhat_limit,
          ess_limit = ess_limit,
          bfmi_limit = bfmi_limit
        )

        full_initial_diagnostic_pass_list[simulation] <-
          full_diagnostics$diagnostic_pass

        null_initial_diagnostic_pass_list[simulation] <-
          null_diagnostics$diagnostic_pass

        # Retry failed models using the same simulated dataset.
        if (
          retry_failed_fits &&
          !isTRUE(full_diagnostics$diagnostic_pass)
        ) {
          full_retry_list[simulation] <- TRUE

          fit_full <- update(
            fit_full,
            newdata = sim_data,
            recompile = FALSE,
            chains = chains,
            cores = 1,
            iter = retry_iter,
            warmup = retry_warmup,
            seed = 40260405 + row * 100000 + simulation,
            control = list(
              adapt_delta = 0.95,
              max_treedepth = 12
            ),
            silent = 2,
            refresh = 0
          )

          full_diagnostics <- diagnose_brms_fit(
            fit = fit_full,
            chains = chains,
            max_treedepth = 12,
            true_slope_sd = slope_sd,
            slope_sd_collapse_threshold = slope_sd_collapse_threshold,
            rhat_limit = rhat_limit,
            ess_limit = ess_limit,
            bfmi_limit = bfmi_limit
          )
        }

        if (
          retry_failed_fits &&
          !isTRUE(null_diagnostics$diagnostic_pass)
        ) {
          null_retry_list[simulation] <- TRUE

          fit_null <- update(
            fit_null,
            newdata = sim_data,
            recompile = FALSE,
            chains = chains,
            cores = 1,
            iter = retry_iter,
            warmup = retry_warmup,
            seed = 50260405 + row * 100000 + simulation,
            control = list(
              adapt_delta = 0.95,
              max_treedepth = 12
            ),
            silent = 2,
            refresh = 0
          )

          null_diagnostics <- diagnose_brms_fit(
            fit = fit_null,
            chains = chains,
            max_treedepth = 12,
            true_slope_sd = slope_sd,
            slope_sd_collapse_threshold = slope_sd_collapse_threshold,
            rhat_limit = rhat_limit,
            ess_limit = ess_limit,
            bfmi_limit = bfmi_limit
          )
        }
        
        full_diagnostic_list[[simulation]] <-
          full_diagnostics
        
        null_diagnostic_list[[simulation]] <-
          null_diagnostics
        
        beta_posterior_mean_list[simulation] <-
          brms::fixef(fit_full)["edi_within", "Estimate"]
        
        sampling_diagnostics_pass <-
          full_diagnostics$diagnostic_pass &&
          null_diagnostics$diagnostic_pass
        
        if (sampling_diagnostics_pass) {
        
          # Repeat bridge sampling to check numerical stability.
          bridge_full_1 <- brms::bridge_sampler(
            fit_full,
            silent = TRUE
          )
        
          bridge_null_1 <- brms::bridge_sampler(
            fit_null,
            silent = TRUE
          )
        
          bridge_full_2 <- brms::bridge_sampler(
            fit_full,
            silent = TRUE
          )
        
          bridge_null_2 <- brms::bridge_sampler(
            fit_null,
            silent = TRUE
          )
        
          log_bf_1 <-
            bridge_full_1$logml -
            bridge_null_1$logml
        
          log_bf_2 <-
            bridge_full_2$logml -
            bridge_null_2$logml
        
          log_bf_list[simulation] <- log_bf_1
          bf_list[simulation] <- exp(log_bf_1)
        
          bridge_logbf_difference_list[simulation] <-
            abs(log_bf_1 - log_bf_2)
        
          # Does repetition change the BF >= 10 decision?
          bridge_decision_stable_list[simulation] <-
            is.finite(log_bf_1) &&
            is.finite(log_bf_2) &&
            (
              (log_bf_1 >= log(bf_threshold)) ==
                (log_bf_2 >= log(bf_threshold))
            )
        }

        p(sprintf(
          "row %d / %d, simulation %d/%d",
          row,
          nrow(informed_parameters),
          simulation,
          num_simulations
        ))
      }
      
      full_diagnostics_all <-
        bind_rows(full_diagnostic_list)
      
      null_diagnostics_all <-
        bind_rows(null_diagnostic_list)
      
      sampling_pass_list <-
        full_diagnostics_all$diagnostic_pass &
        null_diagnostics_all$diagnostic_pass
      
      valid_simulation_list <-
        sampling_pass_list &
        !is.na(bridge_decision_stable_list) &
        bridge_decision_stable_list &
        is.finite(log_bf_list) &
        is.finite(beta_posterior_mean_list)
      
      valid_simulation_list[
        is.na(valid_simulation_list)
      ] <- FALSE
      
      strong_negative_evidence_list <-
        valid_simulation_list &
        log_bf_list >= log(bf_threshold) &
        beta_posterior_mean_list < 0
      
      # Invalid simulations count as unsuccessful in the operational rate.
      strong_negative_evidence_rate <-
        mean(strong_negative_evidence_list)
      
      valid_simulation_rate <-
        mean(valid_simulation_list)
      
      # Evidence rate among diagnostically valid simulations only
      conditional_evidence_rate <-
        if (sum(valid_simulation_list) > 0) {
          mean(
            log_bf_list[valid_simulation_list] >=
              log(bf_threshold) &
              beta_posterior_mean_list[
                valid_simulation_list
              ] < 0
          )
        } else {
          NA_real_
        }
      
      bridge_attempt_rate <-
        mean(!is.na(bridge_decision_stable_list))
      
      bridge_decision_stable_rate <-
        if (any(!is.na(bridge_decision_stable_list))) {
          mean(
            bridge_decision_stable_list,
            na.rm = TRUE
          )
        } else {
          NA_real_
        }
      
      median_bridge_logbf_difference <-
        if (any(is.finite(bridge_logbf_difference_list))) {
          median(
            bridge_logbf_difference_list,
            na.rm = TRUE
          )
        } else {
          NA_real_
        }

      # Evaluate recovery among diagnostically valid Full-model fits.
      slope_recovery_diagnostics <-
        full_diagnostics_all |>
        filter(diagnostic_pass)

      n_slope_recovery_valid <-
        nrow(slope_recovery_diagnostics)

      slope_recovery_valid_rate <-
        n_slope_recovery_valid / num_simulations

      if (n_slope_recovery_valid > 0) {
        mean_estimated_slope_sd <- mean(
          slope_recovery_diagnostics$estimated_slope_sd
        )

        median_estimated_slope_sd <- median(
          slope_recovery_diagnostics$estimated_slope_sd
        )

        slope_sd_bias <-
          mean_estimated_slope_sd - slope_sd

        slope_sd_relative_bias <-
          slope_sd_bias / slope_sd

        slope_sd_rmse <- sqrt(
          mean(
            slope_recovery_diagnostics$slope_sd_squared_error
          )
        )

        slope_sd_coverage <- mean(
          slope_recovery_diagnostics$slope_sd_coverage
        )

        mean_slope_sd_interval_width <- mean(
          slope_recovery_diagnostics$slope_sd_interval_width
        )

        median_slope_sd_interval_width <- median(
          slope_recovery_diagnostics$slope_sd_interval_width
        )

        median_slope_sd_interval_width_ratio <-
          median_slope_sd_interval_width / slope_sd

        slope_sd_collapse_rate <- mean(
          slope_recovery_diagnostics$slope_sd_collapsed
        )
      } else {
        mean_estimated_slope_sd <- NA_real_
        median_estimated_slope_sd <- NA_real_
        slope_sd_bias <- NA_real_
        slope_sd_relative_bias <- NA_real_
        slope_sd_rmse <- NA_real_
        slope_sd_coverage <- NA_real_
        mean_slope_sd_interval_width <- NA_real_
        median_slope_sd_interval_width <- NA_real_
        median_slope_sd_interval_width_ratio <- NA_real_
        slope_sd_collapse_rate <- NA_real_
      }

      random_slope_recovery_pass <-
        is.finite(slope_recovery_valid_rate) &&
        is.finite(slope_sd_relative_bias) &&
        is.finite(slope_sd_coverage) &&
        is.finite(median_slope_sd_interval_width_ratio) &&
        is.finite(slope_sd_collapse_rate) &&
        slope_recovery_valid_rate >= recovery_min_valid_rate &&
        abs(slope_sd_relative_bias) <=
          recovery_max_abs_relative_bias &&
        slope_sd_coverage >= recovery_min_coverage &&
        slope_sd_coverage <= recovery_max_coverage &&
        median_slope_sd_interval_width_ratio <=
          recovery_max_median_interval_width_ratio &&
        slope_sd_collapse_rate <= recovery_max_collapse_rate

      full_retry_rescue_rate <-
        if (sum(full_retry_list) > 0) {
          mean(
            full_diagnostics_all$diagnostic_pass[
              full_retry_list
            ]
          )
        } else {
          NA_real_
        }

      null_retry_rescue_rate <-
        if (sum(null_retry_list) > 0) {
          mean(
            null_diagnostics_all$diagnostic_pass[
              null_retry_list
            ]
          )
        } else {
          NA_real_
        }
      
        true_within_slope <-
          informed_parameters$within_slope_mean[row]
        
        tibble::tibble(
          intercept_mean =
            informed_parameters$intercept_mean[row],
        
          within_slope_mean =
            true_within_slope,
        
          between_slope_mean =
            informed_parameters$between_slope_mean[row],
        
          strong_negative_evidence_rate =
            strong_negative_evidence_rate,
        
          conditional_evidence_rate =
            conditional_evidence_rate,
        
          valid_simulation_rate =
            valid_simulation_rate,
        
          full_diagnostic_pass_rate =
            mean(full_diagnostics_all$diagnostic_pass),

          full_initial_diagnostic_pass_rate =
            mean(full_initial_diagnostic_pass_list),

          full_retry_rate =
            mean(full_retry_list),

          full_retry_rescue_rate =
            full_retry_rescue_rate,
        
          null_diagnostic_pass_rate =
            mean(null_diagnostics_all$diagnostic_pass),

          null_initial_diagnostic_pass_rate =
            mean(null_initial_diagnostic_pass_list),

          null_retry_rate =
            mean(null_retry_list),

          null_retry_rescue_rate =
            null_retry_rescue_rate,
        
          median_bridge_logbf_difference =
            median_bridge_logbf_difference,
        
          mean_estimated_slope_sd =
            mean_estimated_slope_sd,

          median_estimated_slope_sd =
            median_estimated_slope_sd,
        
          slope_sd_bias =
            slope_sd_bias,

          slope_sd_relative_bias =
            slope_sd_relative_bias,

          slope_sd_rmse =
            slope_sd_rmse,
        
          slope_sd_coverage =
            slope_sd_coverage,

          mean_slope_sd_interval_width =
            mean_slope_sd_interval_width,

          median_slope_sd_interval_width =
            median_slope_sd_interval_width,

          median_slope_sd_interval_width_ratio =
            median_slope_sd_interval_width_ratio,

          slope_sd_collapse_threshold =
            slope_sd_collapse_threshold,

          slope_sd_collapse_rate =
            slope_sd_collapse_rate,

          slope_recovery_valid_rate =
            slope_recovery_valid_rate,

          random_slope_recovery_pass =
            random_slope_recovery_pass,
        
          total_divergences =
            sum(
              full_diagnostics_all$n_divergent,
              null_diagnostics_all$n_divergent
            ),
        
          TPR = ifelse(
            true_within_slope < 0,
            strong_negative_evidence_rate,
            NA_real_
          ),
        
          FPR = ifelse(
            true_within_slope == 0,
            strong_negative_evidence_rate,
            NA_real_
          ),
          
          full_divergence_pass_rate =
            mean(full_diagnostics_all$divergence_pass),
          
          full_rhat_pass_rate =
            mean(full_diagnostics_all$rhat_pass),
          
          full_bulk_ess_pass_rate =
            mean(full_diagnostics_all$bulk_ess_pass),
          
          full_tail_ess_pass_rate =
            mean(full_diagnostics_all$tail_ess_pass),
          
          full_bfmi_pass_rate =
            mean(full_diagnostics_all$bfmi_pass),
          
          full_treedepth_pass_rate =
            mean(full_diagnostics_all$treedepth_pass),
          
          null_divergence_pass_rate =
            mean(null_diagnostics_all$divergence_pass),
          
          null_rhat_pass_rate =
            mean(null_diagnostics_all$rhat_pass),
          
          null_bulk_ess_pass_rate =
            mean(null_diagnostics_all$bulk_ess_pass),
          
          null_tail_ess_pass_rate =
            mean(null_diagnostics_all$tail_ess_pass),
          
          null_bfmi_pass_rate =
            mean(null_diagnostics_all$bfmi_pass),
          
          null_treedepth_pass_rate =
            mean(null_diagnostics_all$treedepth_pass),
          
          bridge_attempt_rate =
            bridge_attempt_rate,
          
          bridge_decision_stable_rate =
            bridge_decision_stable_rate
        )
    },
    .options = furrr_options(
      seed = 20260405,
      packages = c(
        "brms",
        "bridgesampling",
        "MASS",
        "dplyr",
        "tibble"
      )
    )
  )
})

head(bfda_simulated_data_parallel)

# Summarise retries and random-slope recovery by generating condition.
bfda_simulated_data_parallel |>
  select(
    within_slope_mean,
    null_initial_diagnostic_pass_rate,
    null_retry_rate,
    null_retry_rescue_rate,
    null_diagnostic_pass_rate,
    slope_recovery_valid_rate,
    mean_estimated_slope_sd,
    slope_sd_relative_bias,
    slope_sd_rmse,
    slope_sd_coverage,
    median_slope_sd_interval_width,
    median_slope_sd_interval_width_ratio,
    slope_sd_collapse_rate,
    random_slope_recovery_pass
  ) |>
  arrange(within_slope_mean)

# Summarise whether recovery criteria pass across all conditions.
three_session_random_slope_assessment <-
  bfda_simulated_data_parallel |>
  summarise(
    all_conditions_pass =
      all(random_slope_recovery_pass),
    minimum_recovery_valid_rate =
      min(slope_recovery_valid_rate),
    maximum_absolute_relative_bias =
      max(abs(slope_sd_relative_bias)),
    minimum_coverage =
      min(slope_sd_coverage),
    maximum_collapse_rate =
      max(slope_sd_collapse_rate),
    median_interval_width_across_conditions =
      median(median_slope_sd_interval_width),
    maximum_median_interval_width_ratio =
      max(median_slope_sd_interval_width_ratio)
  )

three_session_random_slope_assessment

end_time <- Sys.time()
start_time
end_time - start_time
```

The main output columns are:

- `null_initial_diagnostic_pass_rate` is the success rate with the original 4,000-iteration setting.
- `null_retry_rate` is the proportion that required the longer 8,000-iteration refit.
- `null_retry_rescue_rate` is the proportion of retried Null fits that passed after the longer run.
- `null_diagnostic_pass_rate` is the final rate after retry and is the rate used for the Bayes-factor analysis.
- `random_slope_recovery_pass` evaluates the three-observation design for each population-slope condition using the pre-specified valid-fit, relative-bias, coverage, interval-width, and near-zero-collapse criteria.
- `all_conditions_pass` is `TRUE` only when random-slope recovery passes under every simulated population-slope condition. The component metrics should be reported alongside this binary summary.

# 5. Save results

```{r save-results}
dir.create(
  "results",
  showWarnings = FALSE,
  recursive = TRUE
)

write.csv(
  bfda_simulated_data_parallel,
  sprintf(
    "results/bfda_simulated_data_parallel_brms_resid_%s_nsim_%d.csv",
    format(residual_sd, trim = TRUE),
    num_simulations
  ),
  row.names = FALSE
)

write.csv(
  three_session_random_slope_assessment,
  "results/three_session_random_slope_assessment.csv",
  row.names = FALSE
)
```

# 6. Plot the evidence rate

```{r plot-evidence-rate}
ggplot(
  bfda_simulated_data_parallel,
  aes(
    x = within_slope_mean,
    y = strong_negative_evidence_rate,
    group = factor(intercept_mean),
    colour = factor(intercept_mean)
  )
) +
  geom_line() +
  geom_point() +
  labs(
    title = "True positive rate (TPR) for the within-participant daytime mEDI effect",
    x = "Population within-participant mEDI slope",
    y = "TPR",
    colour = "Intercept mean"
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )
```
