---
title: "bfda_recruitment_missingness_sensitivity_DLE-SENSE"
author: "Taisuke Eto"
date: "2026-08-07"
output: html_document
---

This script evaluates the planned recruitment rule under incomplete data. 
The simulation stops when either 50 participants have three valid observations or 60 participants have been enrolled.

Participants with two valid observations are retained in CH1. 
Those with fewer than two are excluded before the within- and between-person decomposition. 
The stopping rule is based only on data completeness.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Packages and analysis settings

```{r packages-and-settings}
library(future)
library(furrr)
library(brms)
library(bridgesampling)
library(MASS)
library(tidyverse)
library(progressr)
library(posterior)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1"
)
rstan::rstan_options(auto_write = TRUE)

# Recruitment settings
target_complete <- 50
maximum_enrolment <- 60

# Effect conditions
within_slopes <- c(-0.3, 0)
intercept_mean <- 0
between_slope_mean <- 0

# Completeness scenarios: probabilities for three, two, or one valid observation
completeness_scenarios <- list(
  high = c(`3` = 0.90, `2` = 0.07, `1` = 0.03),
  expected = c(`3` = 0.85, `2` = 0.10, `1` = 0.05),
  stress = c(`3` = 0.70, `2` = 0.20, `1` = 0.10)
)

num_simulations <- 100  # use a small value for testing
bf_threshold <- 10

intercept_sd <- 0.2
slope_sd <- 0.2
residual_sd <- 0.6
intercept_slope_cor <- 0

chains <- 4
iter <- 4000
warmup <- 1000
retry_failed_fits <- TRUE
retry_iter <- 8000
retry_warmup <- 2000

rhat_limit <- 1.01
ess_limit <- 100 * chains
bfmi_limit <- 0.30
max_treedepth <- 12

workers <- min(
  16,
  length(within_slopes) * length(completeness_scenarios),
  max(1, parallel::detectCores(logical = TRUE) - 1)
)
plan(multisession, workers = workers)
```

# 2. Predictor data and simulation conditions

Observed mEDI profiles are used as the source distribution for the simulated predictor values.

```{r predictor-data-and-design-grid}
medi_dose_data <- read.csv("medi_data_standardised.csv") |>
  filter(is.finite(log_mean_dose_standardised))

medi_values_by_participant <- medi_dose_data |>
  group_by(Id) |>
  summarise(values = list(log_mean_dose_standardised), .groups = "drop") |>
  filter(lengths(values) > 0) |>
  pull(values)

stopifnot(length(medi_values_by_participant) > 0)

normalise_probabilities <- function(probabilities) {
  stopifnot(
    !is.null(names(probabilities)),
    all(names(probabilities) %in% c("0", "1", "2", "3")),
    all(is.finite(probabilities)),
    all(probabilities >= 0),
    sum(probabilities) > 0
  )
  probabilities / sum(probabilities)
}

design_grid <- tidyr::crossing(
  scenario = names(completeness_scenarios),
  within_slope_mean = within_slopes
) |>
  mutate(
    intercept_mean = intercept_mean,
    between_slope_mean = between_slope_mean
  )

# Apply the recruitment stopping rule.
simulate_recruitment <- function(probabilities) {
  probabilities <- normalise_probabilities(probabilities)
  valid_n <- integer(maximum_enrolment)
  enrolled_n <- 0
  complete_n <- 0

  while (
    enrolled_n < maximum_enrolment &&
    complete_n < target_complete
  ) {
    enrolled_n <- enrolled_n + 1
    valid_n[enrolled_n] <- sample(
      as.integer(names(probabilities)),
      size = 1,
      prob = probabilities
    )
    complete_n <- complete_n + (valid_n[enrolled_n] == 3)
  }

  tibble(
    enrolled_id = seq_len(enrolled_n),
    n_valid = valid_n[seq_len(enrolled_n)]
  )
}

# Sample valid sessions and calculate the within- and between-person components.
sample_incomplete_medi_profile <- function(recruitment) {
  analysis_participants <- recruitment |>
    filter(n_valid >= 2) |>
    mutate(id = factor(row_number()))

  if (nrow(analysis_participants) < 2) {
    stop("Too few participants with at least two valid observations.")
  }

  sampled <- purrr::pmap_dfr(
    analysis_participants,
    function(enrolled_id, n_valid, id) {
      donor_values <- medi_values_by_participant[[
        sample.int(length(medi_values_by_participant), 1)
      ]]
      tibble(
        enrolled_id = enrolled_id,
        id = factor(id, levels = levels(analysis_participants$id)),
        session = seq_len(n_valid),
        edi_total = sample(
          donor_values,
          size = n_valid,
          replace = length(donor_values) < n_valid
        )
      )
    }
  ) |>
    group_by(id) |>
    mutate(
      edi_between_raw = mean(edi_total),
      edi_within_raw = edi_total - edi_between_raw
    ) |>
    ungroup()

  between_values <- sampled |>
    distinct(id, edi_between_raw) |>
    pull(edi_between_raw)

  between_mean <- mean(between_values)
  between_sd <- sd(between_values)
  within_sd <- sd(sampled$edi_within_raw)

  if (!all(is.finite(c(between_sd, within_sd))) ||
      between_sd == 0 || within_sd == 0) {
    stop("Unable to standardise simulated mEDI components.")
  }

  sampled |>
    mutate(
      edi_between = (edi_between_raw - between_mean) / between_sd,
      edi_within = edi_within_raw / within_sd,
      id_numeric = as.integer(id)
    )
}
```

# 3. Priors and model diagnostics

The Full model includes the population-level within-person effect; the Null model omits this effect while retaining the same random-effects structure.

```{r priors-and-diagnostics}
full_priors <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(student_t(1, 0, sqrt(2) / 4), class = "b", coef = "edi_within"),
  prior(student_t(1, 0, sqrt(2) / 4), class = "b", coef = "edi_between"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor"),
  prior(student_t(3, 0, 0.5), class = "sigma")
)

null_priors <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(student_t(1, 0, sqrt(2) / 4), class = "b", coef = "edi_between"),
  prior(student_t(3, 0, 0.5), class = "sd"),
  prior(lkj(2), class = "cor"),
  prior(student_t(3, 0, 0.5), class = "sigma")
)

random_effect_covariance <- matrix(
  c(
    intercept_sd^2,
    intercept_slope_cor * intercept_sd * slope_sd,
    intercept_slope_cor * intercept_sd * slope_sd,
    slope_sd^2
  ),
  nrow = 2
)

diagnose_fit <- function(fit) {
  sampler <- rstan::get_sampler_params(fit$fit, inc_warmup = FALSE)
  n_divergent <- sum(vapply(sampler, function(x) sum(x[, "divergent__"]), numeric(1)))
  n_treedepth <- sum(vapply(sampler, function(x) sum(x[, "treedepth__"] >= max_treedepth), numeric(1)))
  min_bfmi <- min(rstan::get_bfmi(fit$fit))

  draws <- posterior::as_draws_df(fit)
  variables <- posterior::variables(draws)
  variables <- variables[grepl("^(b_|sd_|cor_|sigma$)", variables)]
  diagnostics <- posterior::summarise_draws(
    posterior::subset_draws(draws, variable = variables),
    "rhat", "ess_bulk", "ess_tail"
  )

  finite <- all(is.finite(diagnostics$rhat)) &&
    all(is.finite(diagnostics$ess_bulk)) &&
    all(is.finite(diagnostics$ess_tail))

  slope_draws <- draws[["sd_id__edi_within"]]
  slope_lower <- unname(quantile(slope_draws, 0.025))
  slope_upper <- unname(quantile(slope_draws, 0.975))

  tibble(
    diagnostic_pass = finite && n_divergent == 0 && n_treedepth == 0 &&
      max(diagnostics$rhat) <= rhat_limit &&
      min(diagnostics$ess_bulk) >= ess_limit &&
      min(diagnostics$ess_tail) >= ess_limit &&
      is.finite(min_bfmi) && min_bfmi >= bfmi_limit,
    n_divergent = n_divergent,
    n_max_treedepth = n_treedepth,
    max_rhat = max(diagnostics$rhat),
    min_bulk_ess = min(diagnostics$ess_bulk),
    min_tail_ess = min(diagnostics$ess_tail),
    min_bfmi = min_bfmi,
    estimated_slope_sd = mean(slope_draws),
    slope_sd_coverage = slope_lower <= slope_sd && slope_upper >= slope_sd,
    slope_sd_interval_width = slope_upper - slope_lower
  )
}

fit_or_update <- function(previous_fit, formula, priors, data, seed, iterations, warmup_n) {
  if (is.null(previous_fit)) {
    brm(
      formula, data = data, family = gaussian(), prior = priors,
      backend = "rstan", chains = chains, cores = 1,
      iter = iterations, warmup = warmup_n, seed = seed,
      save_pars = save_pars(all = TRUE),
      control = list(adapt_delta = 0.95, max_treedepth = max_treedepth),
      silent = 2, refresh = 0
    )
  } else {
    update(
      previous_fit, newdata = data, recompile = FALSE,
      chains = chains, cores = 1, iter = iterations, warmup = warmup_n,
      seed = seed,
      control = list(adapt_delta = 0.95, max_treedepth = max_treedepth),
      silent = 2, refresh = 0
    )
  }
}
```

# 4. Recruitment and BFDA simulations

For each condition, the script simulates recruitment, generates the outcome, fits the Full and Null models, checks model diagnostics, and calculates the Bayes factor. 
Failed fits are retried with additional iterations.

```{r run-simulations}
handlers("progress")
start_time <- Sys.time()

with_progress({
  p <- progressor(steps = nrow(design_grid) * num_simulations)

  simulation_level_results <- future_map_dfr(
    seq_len(nrow(design_grid)),
    function(grid_row) {
      condition <- design_grid[grid_row, ]
      probabilities <- completeness_scenarios[[condition$scenario]]
      # Reuse compiled models within each condition.
      fit_cache <- new.env(parent = emptyenv())
      fit_cache$full <- NULL
      fit_cache$null <- NULL

      map_dfr(seq_len(num_simulations), function(simulation) {
        recruitment <- simulate_recruitment(probabilities)
        sim_data <- sample_incomplete_medi_profile(recruitment)
        n_analysis <- n_distinct(sim_data$id)

        random_effects <- MASS::mvrnorm(
          n = n_analysis,
          mu = c(0, 0),
          Sigma = random_effect_covariance
        )

        sim_data <- sim_data |>
          mutate(
            y = condition$intercept_mean +
              random_effects[id_numeric, 1] +
              condition$between_slope_mean * edi_between +
              (condition$within_slope_mean + random_effects[id_numeric, 2]) * edi_within +
              rnorm(n(), 0, residual_sd)
          )

        base_seed <- 20260807L + grid_row * 100000L + simulation
        fit_cache$full <- fit_or_update(
          fit_cache$full,
          y ~ edi_within + edi_between + (1 + edi_within | id),
          full_priors, sim_data, base_seed, iter, warmup
        )
        fit_cache$null <- fit_or_update(
          fit_cache$null,
          y ~ edi_between + (1 + edi_within | id),
          null_priors, sim_data, base_seed + 10000000L, iter, warmup
        )

        full_diag <- diagnose_fit(fit_cache$full)
        null_diag <- diagnose_fit(fit_cache$null)
        full_retry <- FALSE
        null_retry <- FALSE

        if (retry_failed_fits && !isTRUE(full_diag$diagnostic_pass)) {
          full_retry <- TRUE
          fit_cache$full <- fit_or_update(
            fit_cache$full,
            y ~ edi_within + edi_between + (1 + edi_within | id),
            full_priors, sim_data, base_seed + 20000000L,
            retry_iter, retry_warmup
          )
          full_diag <- diagnose_fit(fit_cache$full)
        }
        if (retry_failed_fits && !isTRUE(null_diag$diagnostic_pass)) {
          null_retry <- TRUE
          fit_cache$null <- fit_or_update(
            fit_cache$null,
            y ~ edi_between + (1 + edi_within | id),
            null_priors, sim_data, base_seed + 30000000L,
            retry_iter, retry_warmup
          )
          null_diag <- diagnose_fit(fit_cache$null)
        }

        diagnostics_pass <- full_diag$diagnostic_pass && null_diag$diagnostic_pass
        log_bf_1 <- log_bf_2 <- NA_real_

        # Evaluate recovery of the population-level within-person slope.
        full_draws <- posterior::as_draws_df(fit_cache$full)
        beta_draws <- full_draws[["b_edi_within"]]
        beta_mean <- mean(beta_draws)
        beta_lower <- unname(quantile(beta_draws, 0.025))
        beta_upper <- unname(quantile(beta_draws, 0.975))
        beta_error <- beta_mean - condition$within_slope_mean
        beta_squared_error <- beta_error^2
        beta_coverage <-
          beta_lower <= condition$within_slope_mean &&
          beta_upper >= condition$within_slope_mean
        beta_interval_width <- beta_upper - beta_lower

        if (diagnostics_pass) {
          full_bridge_1 <- brms::bridge_sampler(fit_cache$full, silent = TRUE)
          null_bridge_1 <- brms::bridge_sampler(fit_cache$null, silent = TRUE)
          full_bridge_2 <- brms::bridge_sampler(fit_cache$full, silent = TRUE)
          null_bridge_2 <- brms::bridge_sampler(fit_cache$null, silent = TRUE)
          log_bf_1 <- full_bridge_1$logml - null_bridge_1$logml
          log_bf_2 <- full_bridge_2$logml - null_bridge_2$logml
        }

        bridge_stable <- is.finite(log_bf_1) && is.finite(log_bf_2) &&
          ((log_bf_1 >= log(bf_threshold)) == (log_bf_2 >= log(bf_threshold)))
        valid <- diagnostics_pass && bridge_stable && is.finite(beta_mean)
        success <- valid && log_bf_1 >= log(bf_threshold) && beta_mean < 0

        p(sprintf("condition %d/%d, simulation %d/%d", grid_row, nrow(design_grid), simulation, num_simulations))

        tibble(
          scenario = condition$scenario,
          within_slope_mean = condition$within_slope_mean,
          simulation = simulation,
          enrolled_n = nrow(recruitment),
          target_complete_reached = sum(recruitment$n_valid == 3) >= target_complete,
          complete_n = sum(recruitment$n_valid == 3),
          two_observation_n = sum(recruitment$n_valid == 2),
          one_or_zero_observation_n = sum(recruitment$n_valid < 2),
          analysis_n = n_analysis,
          total_observations = nrow(sim_data),
          beta_posterior_mean = beta_mean,
          beta_lower_95 = beta_lower,
          beta_upper_95 = beta_upper,
          beta_error = beta_error,
          beta_squared_error = beta_squared_error,
          beta_coverage_95 = beta_coverage,
          beta_interval_width_95 = beta_interval_width,
          log_bf = log_bf_1,
          bridge_logbf_difference = abs(log_bf_1 - log_bf_2),
          bridge_decision_stable = bridge_stable,
          full_diagnostic_pass = full_diag$diagnostic_pass,
          null_diagnostic_pass = null_diag$diagnostic_pass,
          full_retry = full_retry,
          null_retry = null_retry,
          valid_simulation = valid,
          strong_negative_evidence = success,
          estimated_slope_sd = full_diag$estimated_slope_sd,
          slope_sd_coverage = full_diag$slope_sd_coverage,
          slope_sd_interval_width = full_diag$slope_sd_interval_width
        )
      })
    },
    .options = furrr_options(
      seed = 20260807,
      packages = c("brms", "bridgesampling", "MASS", "tidyverse", "posterior")
    )
  )
})

end_time <- Sys.time()
end_time - start_time
```

# 5. Summarise and save results

Two files are saved: simulation-level results and a summary for each completeness and effect condition.

```{r summarise-and-save}
condition_summary <- simulation_level_results |>
  group_by(scenario, within_slope_mean) |>
  summarise(
    simulations = n(),
    target_complete_reached_rate = mean(target_complete_reached),
    median_enrolled_n = median(enrolled_n),
    median_complete_n = median(complete_n),
    median_analysis_n = median(analysis_n),
    median_total_observations = median(total_observations),
    valid_simulation_rate = mean(valid_simulation),
    full_diagnostic_pass_rate = mean(full_diagnostic_pass),
    null_diagnostic_pass_rate = mean(null_diagnostic_pass),
    bridge_decision_stable_rate = mean(bridge_decision_stable, na.rm = TRUE),
    median_bridge_logbf_difference = median(bridge_logbf_difference, na.rm = TRUE),
    strong_negative_evidence_rate = mean(strong_negative_evidence),
    # Evaluate slope recovery among Full models that pass diagnostics.
    mean_beta_estimate = mean(beta_posterior_mean[full_diagnostic_pass]),
    beta_bias = mean(beta_error[full_diagnostic_pass]),
    beta_rmse = sqrt(mean(beta_squared_error[full_diagnostic_pass])),
    beta_coverage_rate_95 = mean(beta_coverage_95[full_diagnostic_pass]),
    mean_beta_interval_width_95 = mean(beta_interval_width_95[full_diagnostic_pass]),
    mean_estimated_slope_sd = mean(estimated_slope_sd[full_diagnostic_pass]),
    slope_sd_relative_bias = (mean_estimated_slope_sd - slope_sd) / slope_sd,
    slope_sd_coverage = mean(slope_sd_coverage[full_diagnostic_pass]),
    median_slope_sd_interval_width = median(slope_sd_interval_width[full_diagnostic_pass]),
    .groups = "drop"
  ) |>
  mutate(
    TPR = if_else(within_slope_mean < 0, strong_negative_evidence_rate, NA_real_),
    FPR = if_else(within_slope_mean == 0, strong_negative_evidence_rate, NA_real_)
  )

condition_summary

dir.create("results", showWarnings = FALSE, recursive = TRUE)
write.csv(
  simulation_level_results,
  sprintf("results/bfda_recruitment_missingness_simulation_level_nsim_%d.csv", num_simulations),
  row.names = FALSE
)
write.csv(
  condition_summary,
  sprintf("results/bfda_recruitment_missingness_summary_nsim_%d.csv", num_simulations),
  row.names = FALSE
)
```

# 6. Plot evidence rates

```{r plot}
ggplot(
  condition_summary,
  aes(x = scenario, y = strong_negative_evidence_rate, fill = factor(within_slope_mean))
) +
  geom_col(position = "dodge") +
  labs(
    title = "Evidence rate under recruitment and missingness scenarios",
    x = "Completeness scenario",
    y = "BF10 >= 10 and posterior mean < 0",
    fill = "True within slope"
  ) +
  theme_bw()
```

