---
title: "informed_data_simulation_DLE-SENSE"
author: "Taisuke Eto"
date: "2026-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(melidosData)
```

# Overview

This script identifies plausible parameter combinations for the Bayesian factor design analysis (BFDA). 
It uses melatonin suppression data from a previous study and daytime melanopic EDI data from the MeliDos dataset.

The simulated model separates daytime melanopic EDI into within-person and
between-person components:

`outcome ~ edi_within + edi_between + (1 + edi_within | participant)`

The script:

1. prepares and standardises the reference outcome and exposure data;
2. defines candidate fixed- and random-effect parameters;
3. simulates repeated-measures datasets;
4. compares their means and SDs with bootstrap reference intervals; and
5. saves plausible parameter combinations for the BFDA.

The reference outcome dataset has one observation per participant. 
It therefore cannot identify random-effect and residual SDs separately. 
These quantities are screened over prespecified candidate values rather than estimated directly.

# 1. Prepare reference data

## Melatonin suppression and daytime melanopic EDI

```{r import-and-prepare-reference-data}
# Reference melatonin suppression data
mlt_data <- read.csv("VR_paper_melatonin_results.csv")

# Summarise the unstandardised outcome
mlt_raw_data_summary <- mlt_data |>
  summarise(
    mean   = mean(mel_supp),
    sd     = sd(mel_supp),
    median = median(mel_supp),
    iqr    = IQR(mel_supp),
    lower_bound     = quantile(mel_supp, 0.025),
    upper_bound     = quantile(mel_supp, 0.975)
  )


# MeliDos light-exposure data
medi_data_raw <- load_data(
  site = c("BAUA", "MPI", "TUM")
)
medi_data_de <- bind_rows(
  medi_data_raw$BAUA,
  medi_data_raw$MPI,
  medi_data_raw$TUM
)


# Sleep diaries define the daytime window
sleep_data_raw <- load_data(
  "sleepdiaries", site = c("BAUA", "MPI", "TUM")
)
sleep_data_de <- bind_rows(
  sleep_data_raw$BAUA,
  sleep_data_raw$MPI,
  sleep_data_raw$TUM
)

daytime_window <- sleep_data_de |>
  arrange(Id, wake) |>
  group_by(Id) |>
  mutate(
    window_start = wake,
    window_end = lead(bedtime) - hours(7),
    window_start_date = as_date(window_start),
    window_end_date = as_date(window_end)
  ) |>
  ungroup() |>
  filter(
    !is.na(window_start),
    !is.na(window_end),
    window_end > window_start,
    window_start_date == window_end_date
  ) |>
  select(Id, bedtime, wake, window_start, window_end, window_start_date)


# Match light exposure to the corresponding daytime window
medi_sleep_data_de <- medi_data_de |>
  mutate(window_start_date = as_date(Datetime)) |>
  left_join(daytime_window, by = c("Id", "window_start_date"))


# Sum exposure from wake time to seven hours before bedtime
daytime_medi_de <- medi_sleep_data_de |>
  filter(!is.na(MEDI)) |>
  mutate(
    date = window_start_date,
    time = hms::as_hms(Datetime),
    window_start_time = hms::as_hms(window_start),
    window_end_time = hms::as_hms(window_end)
  ) |>
  filter(
    time >= window_start_time,
    time <= window_end_time
  ) |>
  group_by(Id, date) |>
  summarise(
    daytime_medi_dose = sum(MEDI) * (10 / 3600),
    n_valid = n(),
    .groups = "drop"
  ) |>
  mutate(log_mean_dose = log10(daytime_medi_dose + 0.1))

# Summarise between- and within-person exposure variability
subject_daytime_stats <- daytime_medi_de |>
  group_by(Id) |>
  summarise(
    mean_dose = mean(daytime_medi_dose, na.rm = TRUE),
    sd_dose = sd(daytime_medi_dose, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    log_mean_dose_subject = log10(mean_dose + 0.1)
  )

between_daytime_medi_summary <- subject_daytime_stats |>
  summarise(
    mean = mean(mean_dose),
    sd = sd(mean_dose),
    min = min(mean_dose),
    max = max(mean_dose)
  )

within_daytime_medi_summary <- subject_daytime_stats |>
  summarise(
    mean_sd = mean(sd_dose, na.rm = TRUE),
    sd_sd = sd(sd_dose, na.rm = TRUE),
    median_sd = median(sd_dose, na.rm = TRUE),
    iqr_sd = IQR(sd_dose, na.rm = TRUE)
  )


mlt_raw_data_summary
between_daytime_medi_summary
within_daytime_medi_summary
```

## Standardise outcome and exposure

```{r standardise-reference-data}
mlt_data_standardised <- mlt_data |>
  mutate(
    mel_supp_standardised =
      (mel_supp - mean(mel_supp)) / sd(mel_supp)
  )

mlt_data_standardised_summary <- mlt_data_standardised |>
  summarise(
    mean   = mean(mel_supp_standardised),
    sd     = sd(mel_supp_standardised),
    median = median(mel_supp_standardised),
    lower_bound     = quantile(mel_supp_standardised, 0.025),
    upper_bound     = quantile(mel_supp_standardised, 0.975)
  )

medi_data_standardised <- daytime_medi_de |>
  filter(is.finite(log_mean_dose)) |>
  mutate(
    log_mean_dose_standardised =
      (log_mean_dose - mean(log_mean_dose)) / sd(log_mean_dose)
  )

medi_data_standardised_summary <- medi_data_standardised |>
  summarise(
    mean   = mean(log_mean_dose_standardised),
    sd     = sd(log_mean_dose_standardised),
    median = median(log_mean_dose_standardised),
    lower_bound     = quantile(log_mean_dose_standardised, 0.025),
    upper_bound     = quantile(log_mean_dose_standardised, 0.975)
  )

medi_data_standardised_summary
mlt_data_standardised_summary
```

# 2. Define candidate parameters

We assign a wide range of possible values to the intercept and slope. 
We also screen candidate values for the random-intercept, random-slope, and residual SDs, as well as the correlation between the random intercept and random slope.
These values are assumptions to be checked, rather than variance components estimated from the single-observation reference melatonin dataset.

```{r candidate-parameters}
# Population intercept values on the standardised scale
intercept_mean_list <- round(seq(-0.9, 0.9, by = 0.3), 2)

# Population-level within-person slopes on the standardised scale
within_slope_mean_list <- seq(-0.5, 0.5, by = 0.1)

# The between-person effect is fixed at zero for parameter screening
between_slope_mean <- 0

# Candidate participant-level and observation-level SDs
intercept_sd_list <- c(0.1, 0.2, 0.3)
slope_sd_list <- c(0.1, 0.2, 0.3)
noise_sd_list <- c(0.3, 0.6, 0.9, 1.0)

# Candidate random intercept-slope correlations
intercept_slope_correlation_list <- c(-0.5, 0, 0.5)

# Prospective study structure
n_participants <- 50
n_sessions <- 3

# Number of complete simulated datasets per parameter combination
num_simulations <- 100

# Number of bootstrap samples used to define reference intervals
n_bootstrap <- 5000

# Minimum proportion meeting both reference criteria
plausibility_threshold <- 0.80
```

# 3. Define bootstrap reference intervals

Bootstrap 95% intervals are calculated for the mean and SD of the standardised reference outcome.

```{r bootstrap-reference-intervals}
set.seed(20260731)

reference_z <- mlt_data_standardised$mel_supp_standardised
n_reference <- length(reference_z)

reference_bootstrap <- replicate(
  n_bootstrap,
  {
    bootstrap_sample <- sample(
      reference_z,
      size = n_reference,
      replace = TRUE
    )

    c(
      mean = mean(bootstrap_sample),
      sd = sd(bootstrap_sample)
    )
  }
) |>
  t() |>
  as.data.frame()

reference_intervals <- reference_bootstrap |>
  summarise(
    mean_lower = quantile(mean, 0.025),
    mean_upper = quantile(mean, 0.975),
    sd_lower = quantile(sd, 0.025),
    sd_upper = quantile(sd, 0.975)
  )

reference_intervals
```

# 4. Define simulation functions

Predictor values are sampled from actual MeliDos participants. 
For each simulated participant, one MeliDos participant is sampled as a donor and three days are drawn from that donor. 
This retains more of the observed within- and between-participant exposure structure than drawing from a uniform distribution.

```{r simulation-functions}
medi_values_by_participant <- split(
  medi_data_standardised$log_mean_dose_standardised,
  medi_data_standardised$Id
)

medi_values_by_participant <- medi_values_by_participant[
  lengths(medi_values_by_participant) > 0
]

sample_medi_profile <- function(n_participants, n_sessions) {

  donor_index <- sample.int(
    length(medi_values_by_participant),
    size = n_participants,
    replace = TRUE
  )

  sampled_values <- map(
    donor_index,
    function(index) {

      donor_values <- medi_values_by_participant[[index]]

      donor_values[
        sample.int(
          length(donor_values),
          size = n_sessions,
          replace = length(donor_values) < n_sessions
        )
      ]
    }
  )

  sampled_data <- tibble(
    participant = factor(
      rep(seq_len(n_participants), each = n_sessions)
    ),
    session = rep(
      seq_len(n_sessions),
      times = n_participants
    ),
    edi_total = unlist(
      sampled_values,
      use.names = FALSE
    )
  ) |>
    group_by(participant) |>
    mutate(
      # Participant mean across simulated bright-light sessions
      edi_between_raw = mean(edi_total),

      # Session-specific deviation from that participant's mean
      edi_within_raw = edi_total - edi_between_raw
    ) |>
    ungroup()

  # Standardise within- and between-person components separately
  between_values <- sampled_data |>
    distinct(participant, edi_between_raw) |>
    pull(edi_between_raw)

  between_mean <- mean(between_values)
  between_sd <- sd(between_values)
  within_sd <- sd(sampled_data$edi_within_raw)

  if (
    !is.finite(between_sd) ||
    !is.finite(within_sd) ||
    between_sd == 0 ||
    within_sd == 0
  ) {
    stop("Unable to standardise the sampled within/between mEDI components.")
  }

  sampled_data |>
    mutate(
      edi_between =
        (edi_between_raw - between_mean) / between_sd,

      edi_within =
        edi_within_raw / within_sd
    )
}


prepare_simulation_input <- function(
    n_participants,
    n_sessions) {

  predictor_data <- sample_medi_profile(
    n_participants = n_participants,
    n_sessions = n_sessions
  )

  list(
    predictor_data = predictor_data,
    z_intercept = rnorm(n_participants),
    z_slope = rnorm(n_participants),
    z_noise = rnorm(nrow(predictor_data))
  )
}


simulate_dataset <- function(
    simulation_input,
    intercept_mean,
    within_slope_mean,
    between_slope_mean,
    intercept_sd,
    slope_sd,
    intercept_slope_correlation,
    noise_sd) {

  simulated_data <- simulation_input$predictor_data
  n_participants <- nlevels(simulated_data$participant)

  # Construct correlated standard-normal random effects
  z_slope_correlated <-
    intercept_slope_correlation * simulation_input$z_intercept +
    sqrt(1 - intercept_slope_correlation^2) *
      simulation_input$z_slope

  # Participant-specific random intercept
  participant_intercept <-
    intercept_mean +
    intercept_sd * simulation_input$z_intercept

  # Participant-specific slope for the within-participant predictor
  participant_within_slope <-
    within_slope_mean +
    slope_sd * z_slope_correlated

  participant_index <- as.integer(
    simulated_data$participant
  )

  simulated_data |>
    mutate(
      intercept =
        participant_intercept[participant_index],

      within_slope =
        participant_within_slope[participant_index],

      noise =
        noise_sd * simulation_input$z_noise,

      y =
        intercept +
        between_slope_mean * edi_between +
        within_slope * edi_within +
        noise
    )
}
```

# 5. Screen parameter combinations

```{r run-parameter-screening}
set.seed(20260405)

simulation_time <- system.time({

# Reuse common random draws across parameter combinations
simulation_inputs <- map(
  seq_len(num_simulations),
  ~ prepare_simulation_input(
    n_participants = n_participants,
    n_sessions = n_sessions
  )
)

simulation_result_list <- list()
result_index <- 1

for (intercept_mean in intercept_mean_list) {
  for (within_slope_mean in within_slope_mean_list) {
    for (intercept_sd in intercept_sd_list) {
      for (slope_sd in slope_sd_list) {
        for (intercept_slope_correlation in
             intercept_slope_correlation_list) {
          for (noise_sd in noise_sd_list) {

            # Store one row for each complete simulated dataset
            combination_results <- vector(
              mode = "list",
              length = num_simulations
            )

            for (simulation in seq_len(num_simulations)) {

              simulated_data <- simulate_dataset(
                simulation_input = simulation_inputs[[simulation]],
                intercept_mean = intercept_mean,
                within_slope_mean = within_slope_mean,
                between_slope_mean = between_slope_mean,
                intercept_sd = intercept_sd,
                slope_sd = slope_sd,
                intercept_slope_correlation =
                  intercept_slope_correlation,
                noise_sd = noise_sd
              )

              combination_results[[simulation]] <- tibble(
                simulation = simulation,
                mean_y = mean(simulated_data$y),
                sd_y = sd(simulated_data$y),
                q025_y = quantile(simulated_data$y, 0.025),
                q975_y = quantile(simulated_data$y, 0.975)
              )
            }

            combination_results <- bind_rows(combination_results) |>
              mutate(
                mean_plausible =
                  mean_y >= reference_intervals$mean_lower &
                  mean_y <= reference_intervals$mean_upper,
                sd_plausible =
                  sd_y >= reference_intervals$sd_lower &
                  sd_y <= reference_intervals$sd_upper,
                jointly_plausible = mean_plausible & sd_plausible
              )

            simulation_result_list[[result_index]] <-
              combination_results |>
              summarise(
                intercept_mean = intercept_mean,
                within_slope_mean = within_slope_mean,
                between_slope_mean = between_slope_mean,
                intercept_sd = intercept_sd,
                slope_sd = slope_sd,
                intercept_slope_correlation =
                  intercept_slope_correlation,
                noise_sd = noise_sd,
                mean_y = mean(mean_y),
                sd_y = mean(sd_y),
                mean_q025_y = mean(q025_y),
                mean_q975_y = mean(q975_y),
                mean_acceptance_rate = mean(mean_plausible),
                sd_acceptance_rate = mean(sd_plausible),
                joint_acceptance_rate = mean(jointly_plausible),
                accepted = if_else(
                  joint_acceptance_rate >= plausibility_threshold,
                  "yes",
                  "no"
                )
              )

            result_index <- result_index + 1
          }
        }
      }
    }
  }
}



simulated_data_results <- bind_rows(simulation_result_list) |>
  mutate(accepted = factor(accepted, levels = c("no", "yes")))

# Keep only accepted parameter combinations
accepted_params <- simulated_data_results |>
  filter(accepted == "yes")

# Summary of accepted parameters
accepted_params_summary <- accepted_params |>
  summarise(
    n_accepted = n(),
    min_intercept = min(intercept_mean),
    max_intercept = max(intercept_mean),
    min_edi = min(within_slope_mean),
    median_edi = median(within_slope_mean),
    max_edi = max(within_slope_mean),
    min_intercept_sd = min(intercept_sd),
    max_intercept_sd = max(intercept_sd),
    min_slope_sd = min(slope_sd),
    max_slope_sd = max(slope_sd),
    min_intercept_slope_correlation =
      min(intercept_slope_correlation),
    max_intercept_slope_correlation =
      max(intercept_slope_correlation),
    min_noise_sd = min(noise_sd),
    max_noise_sd = max(noise_sd)
  )

})

simulation_time

accepted_params_summary
```


# 6. Review screening results

## Fixed effects

```{r plot-fixed-effect-acceptance}
fixed_effect_acceptance_summary <- simulated_data_results |>
  group_by(
    intercept_mean,
    within_slope_mean,
    noise_sd
  ) |>
  summarise(
    proportion_random_effect_structures_accepted =
      mean(accepted == "yes"),
    .groups = "drop"
  )

ggplot(
  fixed_effect_acceptance_summary,
  aes(
    x = factor(intercept_mean),
    y = factor(within_slope_mean),
    fill = proportion_random_effect_structures_accepted
  )
) +
  geom_tile(color = "grey85", linewidth = 0.25) +
  facet_wrap(
    vars(noise_sd),
    ncol = 2,
    labeller = label_both
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkgreen",
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.25),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Acceptance of fixed-effect parameter combinations",
    subtitle = paste0(
      "Cell colour: proportion of 27 random-effect structures accepted; ",
      "joint plausibility threshold = ",
      scales::percent(plausibility_threshold, accuracy = 1)
    ),
    x = "Population intercept mean",
    y = "Population daytime melanopic EDI slope mean",
    fill = "Structures\naccepted"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
  )
```

## Variance components

```{r plot-variance-component-acceptance}
sd_acceptance_summary <- simulated_data_results |>
  group_by(
    intercept_sd,
    slope_sd,
    intercept_slope_correlation,
    noise_sd
  ) |>
  summarise(
    mean_joint_acceptance_rate = mean(joint_acceptance_rate),
    proportion_parameter_sets_accepted = mean(accepted == "yes"),
    .groups = "drop"
  )

ggplot(
  sd_acceptance_summary,
  aes(
    x = factor(noise_sd),
    y = mean_joint_acceptance_rate,
    group = factor(intercept_slope_correlation),
    color = factor(intercept_slope_correlation)
  )
) +
  geom_line() +
  geom_point(size = 2) +
  facet_grid(
    rows = vars(intercept_sd),
    cols = vars(slope_sd),
    labeller = label_both
  ) +
  labs(
    title = "Plausibility of candidate variance components",
    x = "Residual SD",
    y = "Mean joint acceptance rate",
    color = "Intercept–slope correlation"
  ) +
  theme_bw()
```

## Selected variance components

This check makes the consequence of the selected SD values explicit. 

```{r inspect-selected-parameters}
selected_parameter_check <- simulated_data_results |>
  filter(
    intercept_sd == 0.2,
    slope_sd == 0.2,
    intercept_slope_correlation == 0.0,
    noise_sd == 0.6,
    intercept_mean %in% c(-0.9, -0.6, -0.3, 0, 0.3, 0.6, 0.9)
  )

ggplot(
  selected_parameter_check,
  aes(
    x = within_slope_mean,
    y = joint_acceptance_rate,
    group = factor(intercept_mean),
    colour = factor(intercept_mean)
  )
) +
  geom_hline(
    yintercept = plausibility_threshold,
    linetype = "dashed",
    colour = "grey40"
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(
    aes(shape = accepted),
    size = 3
  ) +
  scale_x_continuous(
    breaks = seq(-0.5, 0.5, by = 0.1)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Acceptance rates for the selected simulation parameters",
    subtitle = paste0(
      "Random-intercept SD = 0.2; random-slope SD = 0.2; ",
      "correlation = 0; residual SD = 0.6"
    ),
    x = "Population slope mean",
    y = "Joint acceptance rate",
    colour = "Population\nintercept mean",
    shape = "Accepted"
  ) +
  theme_bw()
```

# 7. Create and save BFDA inputs

```{r create-informed-parameters}
informed_parameters <- accepted_params |> 
  distinct(
    intercept_mean,
    within_slope_mean,
    between_slope_mean
  ) |> 
  arrange(
    intercept_mean,
    within_slope_mean,
    between_slope_mean
  )

informed_parameters
```


The main BFDA input is `informed_parameters.csv`. The complete screening results and standardised mEDI data are also saved for reproducibility.

```{r save-outputs}
write.csv(
  informed_parameters,
  "informed_parameters.csv",
  row.names = FALSE
)

write.csv(
  simulated_data_results,
  "informed_parameter_screening_all_results.csv",
  row.names = FALSE
)

write.csv(
  medi_data_standardised,
  "medi_data_standardised.csv",
  row.names = FALSE
)
```
